---
title: "Uncertainty, Accuracy, Bias"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=T, message=F, warning=F}
knitr::opts_chunk$set(message=F, warning=F, fig.align = "center",  dev='svg')

library(tidyverse) #loads multiple packages (see https://tidyverse.tidyverse.org/)

#core tidyverse packages loaded:
# ggplot2, for data visualisation. https://ggplot2.tidyverse.org/
# dplyr, for data manipulation. https://dplyr.tidyverse.org/
# tidyr, for data tidying. https://tidyr.tidyverse.org/
# readr, for data import. https://readr.tidyverse.org/
# purrr, for functional programming. https://purrr.tidyverse.org/
# tibble, for tibbles, a modern re-imagining of data frames. https://tibble.tidyverse.org/
# stringr, for strings. https://stringr.tidyverse.org/
# forcats, for factors. https://forcats.tidyverse.org/
# lubridate, for date/times. https://lubridate.tidyverse.org/

#also loads the following packages (less frequently used):
# Working with specific types of vectors:
#     hms, for times. https://hms.tidyverse.org/
# Importing other types of data:
#     feather, for sharing with Python and other languages. https://github.com/wesm/feather
#     haven, for SPSS, SAS and Stata files. https://haven.tidyverse.org/
#     httr, for web apis. https://httr.r-lib.org/
#     jsonlite for JSON. https://arxiv.org/abs/1403.2805
#     readxl, for .xls and .xlsx files. https://readxl.tidyverse.org/
#     rvest, for web scraping. https://rvest.tidyverse.org/
#     xml2, for XML. https://xml2.r-lib.org/
# Modelling
#     modelr, for modelling within a pipeline. https://modelr.tidyverse.org/
#     broom, for turning models into tidy data. https://broom.tidymodels.org/

# Use the conflicted package (<http://conflicted.r-lib.org/>) to force all conflicts to become errors


library(cowplot) #for plot_grid()
library(ggbeeswarm) #for geom_beeswarm()

#set theme for graphs
theme_set(
  theme_classic() +
  theme(
    panel.grid.major.y = element_line(), #no vertical lines by default
    text = element_text(family = "Times New Roman"), #default font
    plot.title = element_text(face="bold"), #graphs titles in bolds
    )
  )

max_new_hosp <- 3036 #historical peak of new hospitalizations
dpi_resolution <- 500 #default dpi resolution for our png graphs
max_ICU_beds <- 6937 #historical peak of ICU beds
```

# Prepare data

Load files generated from the jupyter notebooks

```{r load_data}
errors <-
  read_csv('results/mean_error.csv') %>%
  mutate(
    #to have percentages on graphs
    MAE = MAE/100,
    ME = ME/100,
    `Max Error` = `Max Error`/100,
    MAPE = MAPE/100
    ) %>%
  rename(
    #for more explicit column names
    scenario_ID = ...1,
    scenario_type = scenario,
    `Mean Absolute Error (%)` = MAE,
    `Mean Error (%)` = ME,
    `Maximum Error (%)` = `Max Error`,
    `Mean Absolute Percentage Error` = MAPE,
    `Mean Absolute Error (beds)` = `MAE (beds)`,
    `Maximum Error (beds)` = `Max error (beds)`
    ) %>%
  mutate(
    scenario_ID = as.Date(gsub("Scenario: ", "", scenario_ID)),
    scenario_type = case_when(
      scenario_type == "Optimist" ~ "most\noptimistic",
      scenario_type == "Pessimist" ~ "most\npessimistic",
      scenario_type == "Median" ~ "median"
      )
    )

#order scenarios levels for graph order
errors$scenario_type <- factor(
  errors$scenario_type,
  levels = c(
    'most\npessimistic',
    'median',
    'most\noptimistic'
  )
)

uncertainties <- read_csv('results/global_evaluation.csv') %>%
  rename(
    scenario_ID = ...1,
    ) %>%
  select(`Average uncertainty`, `Max uncertainty`, scenario_ID) %>%
  mutate(
    endpoint = case_when(
      grepl("ICU", scenario_ID) ~ "ICU",
      grepl("New hosp.", scenario_ID) ~ "New hosp."
    ),
    scenario_ID = as.Date(gsub("Scenario: ", "", scenario_ID)),
  )

temporal <- read_csv('results/global_evaluation_dates.csv') %>%
  select(-...1) %>%
  rename(
    scenario_ID = Scenario,
    endpoint = `Scenario type`
    ) %>%
  mutate(
    scenario_ID = as.Date(gsub("Scenario: ", "", scenario_ID))
  ) %>%
  #change time point names from days to weeks
  filter(Period!="56 days - 70 days") %>% #to few points for this period
  mutate(
    Period = case_when(
      Period == "0 days - 14 days" ~ "0-2 weeks",
      Period == "14 days - 28 days" ~ "2-4 weeks",
      Period == "28 days - 42 days" ~ "4-6 weeks",
      Period == "42 days - 56 days" ~ "6-8 weeks",
      Period == "" ~ ""
    )
  )
```

functions used in the code: save_graphs (save png and pdf graphs)

```{r code_functions}
f_save_graph_pdf_png <- function(path_name, graph_width, graph_height, dpi_resolution){
  #pdf
  ggsave(
    paste0(path_name, ".pdf"),
    width=graph_width, height=graph_height, bg="white", 
    device = cairo_pdf #devide cairo for Times New Roman font in pdf
    )
  #png
  ggsave(
    paste0(path_name, ".png"),
    width=graph_width, height=graph_height, bg="white", 
    dpi = dpi_resolution
    )
}
```

# Absolute Errors {.tabset .tabset-fade .tabset-pills}


rajouter flèche en légende et sur graph pour expliquer de accurate à less accurate ? ou alors une légère background color vert orange rouge comme sur les figure qualitatives ?

## Absolute error (MAE vs MAPE) {.tabset .tabset-fade}

```{r function_graph_MAPE_vs_MAE}
#graph code adapted from http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/78-perfect-scatter-plots-with-correlation-and-marginal-histograms/

g_MAPE_vs_MAE <- function(df){
  # Main plot
  pmain <- ggplot(temp)+
    geom_point(
      aes(
        y = `Mean Absolute Error (%)`, x = `Mean Absolute Percentage Error`, 
        color = scenario_type, fill = scenario_type
        ),
      alpha = 0.6, stroke=1
      )+
    theme(
      panel.grid.major.x = element_line(),
      legend.position = c(.75, .4),
      legend.background = element_rect(fill=alpha('white', .5), color="grey")
      ) +
    scale_y_continuous(labels = scales::percent) +
    scale_x_continuous(
      labels=scales::percent(c(seq(0, max(temp$`Mean Absolute Percentage Error`), 1), .5)),
      breaks = c(seq(0, max(temp$`Mean Absolute Percentage Error`), 1), .5)
      ) +
    scale_color_viridis_d() +
    labs(
      y="Mean Absolute Error\n(% of historical bed occupancy peak)",
      x="Mean Absolute Percentage Error\n(% of real epidemic activity)",
      color="scenario from\neach report",
      fill="scenario from\neach report"
    ) +
    geom_hline(yintercept = 1, linetype="dashed", alpha=.4) 
  
  # Marginal densities along x axis
  xdens <- axis_canvas(pmain, axis = "x") +
    geom_boxplot(
      data = temp, 
      aes(x = `Mean Absolute Percentage Error`, fill=scenario_type),
      alpha = 0.5, size = 0.2, linewidth=.5, outlier.shape = 4
      ) +
    scale_fill_viridis_d()
  
  # Marginal densities along y axis
  ydens <- axis_canvas(pmain, axis = "y", coord_flip = TRUE) + # Need to set coord_flip = TRUE, if you plan to use coord_flip()
    geom_boxplot(
      data = temp, 
      aes(x = `Mean Absolute Error (%)`, fill = scenario_type),
      alpha = 0.5, size = 0.2, linewidth=.5, outlier.shape = 4
      ) +
    coord_flip() +
    scale_fill_viridis_d() 
  
  return(
    list(pmain, xdens,  ydens)
    )
}
```

### ICU + New Hosp

when scenarios relate to both ICU and new hosp, we take the mean of the 2

```{r MAPE_vs_MAE_ICU_new_hosp_combined}
#when scenarios relate to both ICU and new hosp, we take the mean of the 2
temp <- errors %>%
  select(
    scenario_ID, scenario_type, 
    `Mean Absolute Percentage Error`, `Mean Absolute Error (%)`
    ) %>%
  group_by(scenario_ID, scenario_type) %>%
  summarise(
    across(
      c(`Mean Absolute Percentage Error`, `Mean Absolute Error (%)`), 
      ~mean(.x)
      )
  )

#create plot
p1 <- insert_xaxis_grob(
  g_MAPE_vs_MAE(temp)[[1]], 
  g_MAPE_vs_MAE(temp)[[2]], 
  grid::unit(.2, "null"), position = "top"
  )
p2<- insert_yaxis_grob(
  p1, 
  g_MAPE_vs_MAE(temp)[[3]], 
  grid::unit(.2, "null"), position = "right"
  )
(g1 <- ggdraw(p2))

#save
f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/MAE_vs_MAPE",
  4, 4, dpi_resolution
)

rm(p1, p2, g1)

#same but with no legend (for subsequent graphs combination)
p1_no_legend <- insert_xaxis_grob(
  g_MAPE_vs_MAE(temp)[[1]] + theme(legend.position = "none"), 
  g_MAPE_vs_MAE(temp)[[2]], 
  grid::unit(.2, "null"), position = "top"
  )
p2_no_legend<- insert_yaxis_grob(
  p1_no_legend, 
  g_MAPE_vs_MAE(temp)[[3]], 
  grid::unit(.2, "null"), position = "right"
  )
g1_no_legend <- ggdraw(p2_no_legend)

rm(p1_no_legend, p2_no_legend, temp)
```

### ICU alone

```{r MAPE_vs_MAE_ICU_alone}
temp <- errors %>%
  filter(endpoints=="ICU")

#create plot
p1 <- insert_xaxis_grob(
  g_MAPE_vs_MAE(temp)[[1]], 
  g_MAPE_vs_MAE(temp)[[2]], 
  grid::unit(.2, "null"), position = "top"
  )
p2<- insert_yaxis_grob(
  p1, 
  g_MAPE_vs_MAE(temp)[[3]], 
  grid::unit(.2, "null"), position = "right"
  )
(g1 <- ggdraw(p2))

rm(p1, p2, g1, temp)
```

### new hosp alone

```{r MAPE_vs_MAE_new_hosp_alone}
temp <- errors %>%
  filter(endpoints=="New hosp.")

#create plot
p1 <- insert_xaxis_grob(
  g_MAPE_vs_MAE(temp)[[1]], 
  g_MAPE_vs_MAE(temp)[[2]], 
  grid::unit(.2, "null"), position = "top"
  )
p2<- insert_yaxis_grob(
  p1, 
  g_MAPE_vs_MAE(temp)[[3]], 
  grid::unit(.2, "null"), position = "right"
  )
(g1 <- ggdraw(p2))

rm(p1, p2, g1, temp)
```

## Absolute mean and max error (MAE)

Supprimer max error ? rnommer ICU et new hops.

```{r max_mean_absolute_error_beds}
temp <- errors %>%
  select(
    scenario_ID, 
    scenario_type, 
    endpoints,
    `Maximum\nAbsolute Error`=`Maximum Error (beds)`, 
    `Mean\nAbsolute Error`= `Mean Absolute Error (beds)`
    ) %>%
  gather(
    key=indicator, value=percentage, 
    `Maximum\nAbsolute Error`, `Mean\nAbsolute Error`
    )

hline_dat = data.frame(
  endpoints=c("ICU", "New hosp."),
  beds=c(max_ICU_beds, max_new_hosp)
  )

(g2 <- ggplot(temp) +
    geom_boxplot(
      aes(scenario_type, percentage, fill=scenario_type),
      position = position_nudge(y = 0, x = -.35),
      outlier.shape = NA, width=.5, alpha=.6
      ) +
    geom_quasirandom(
        aes(scenario_type, percentage, color=scenario_type),
        alpha=.6, width=.1,
        ) +
    geom_hline(
      data=hline_dat, 
      aes(yintercept=beds), 
      linetype="dashed", alpha=.4
      ) +
    facet_grid(endpoints~indicator, scales = "free") +
    labs(
      x='', 
      y='error in terms of beds',
      color="scenarios from\neach report",
      fill="scenarios from\neach report",
      subtitle = "horizontal lines: historical bed occupancy peak"
      ) +
    guides(
      fill="none",
      color = guide_legend(override.aes = list(size = 3))
      ) +
    scale_fill_viridis_d() +
    scale_color_viridis_d() +
    scale_y_continuous(
      labels = scales::label_number(drop0trailing = TRUE)
    ) +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.line.x = element_blank()
    ) +
    geom_hline(yintercept = 0)
)

rm(temp, hline_dat)

#save
f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/mean_max_beds_error",
  6, 4, dpi_resolution
)
```

## Both graphs

```{r all_error_graphs}
plot_grid(
  g2, g1_no_legend ,
  rel_widths = c(5.5/9, 3.5/9),
  labels = c("a", "b")
)

#save
f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/errors",
  9, 3.5, dpi_resolution
)

rm(g1_no_legend, g2)
```

# Bias {.tabset .tabset-fade .tabset-pills}

## Error bias (Mean Error) {.tabset .tabset-fade}

when scenarios relate to both ICU and new hosp, we take the mean of the 2

### all period

```{r bias_mean_error}
temp <- errors %>%
  select(scenario_ID, scenario_type, `Mean Error (%)`) %>%
  group_by(scenario_ID, scenario_type) %>%
  summarise(
    `Mean Error (%)` = mean(`Mean Error (%)`)
    )
temp$scenario_type <- fct_recode(
  temp$scenario_type, 
  `most pessimistic scenario\nof report` = "most\npessimistic", 
  `median scenario\nof report` = "median", 
  `most optimistic scenario\nof report` = "most\noptimistic"
  )

(g1 <- ggplot(temp) +
    geom_boxplot(
      aes(scenario_type, `Mean Error (%)`, fill=scenario_type),
      position = position_nudge(y = 0, x = -.3),
      outlier.shape = NA,
      width=.2, alpha=.6
      ) +
    geom_quasirandom(
        aes(scenario_type, `Mean Error (%)`, color=scenario_type),
        alpha=.6, width=.1
        ) +
    scale_y_continuous(
      labels = scales::percent,
      limits=c(-.8, .8),
      breaks = seq(-.75, .75, .25)
      ) +
    labs(
      x='',
      y='Mean Error'
      ) +
    scale_fill_viridis_d() +
    scale_color_viridis_d() +
    theme(legend.position = 'none') +
    geom_hline(yintercept=0, linetype='dashed') +
    annotate(
      'segment', arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
      x = 3.1, y = 0.1, xend = 3.1, yend = .7,  alpha=.5
      ) +
    annotate(
      'segment', arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
      x = 3.1, y = -0.1, xend = 3.1, yend = -.7, alpha=.5
      ) +
    annotate(
      "text", x = 3.15, y = .1, label = "unbiased", hjust=0, 
      fontface = "italic", family = "Times New Roman",
      ) +
    annotate(
      "text", x = 3, y = -.7, label = "too\noptimistic", hjust=1, vjust=0,
      fontface = "italic", family = "Times New Roman",
      ) +
    annotate(
      "text", x = 3, y = .7, label = "too\npessimistic", hjust=1, vjust=1,
      fontface = "italic", family = "Times New Roman",
      )
)


f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/bias_mean_error",
  6, 4, 500
)
```

### by 2 weeks period

TBD

## Self-assessment bias {.tabset .tabset-fade}

changer le format des dates pour + lisible (comme sur qualitative graph)

```{r tweak_reports_labels_for_better_display_on_graph}
#small offset to position adjacent reports labels on x axis
reports_date <-  data.frame(
  positions = unique(as.Date(gsub("Scenario: ", "", errors$scenario_ID))),
  labels = unique(as.Date(gsub("Scenario: ", "", errors$scenario_ID)))
)
#Aug and Jul 2021
reports_date$positions[reports_date$labels=="2021/07/26"] <- reports_date$positions[reports_date$labels=="2021/07/26"] -10
reports_date$positions[reports_date$labels=="2021/08/05"] <- reports_date$positions[reports_date$labels=="2021/08/05"] +10
#Apr and May 2021
reports_date$positions[reports_date$labels=="2021-04-26"] <- reports_date$positions[reports_date$labels=="2021-04-26"] -5
reports_date$positions[reports_date$labels=="2021-05-21"] <- reports_date$positions[reports_date$labels=="2021-05-21"] +5
#Jan and Feb 2021
reports_date$positions[reports_date$labels=="2021-01-16"] <- reports_date$positions[reports_date$labels=="2021-01-16"] -20
reports_date$positions[reports_date$labels=="2021-02-02"] <- reports_date$positions[reports_date$labels=="2021-02-02"] -10
reports_date$positions[reports_date$labels=="2021-02-08"] <- reports_date$positions[reports_date$labels=="2021-02-08"] +5
reports_date$positions[reports_date$labels=="2021-02-23"] <- reports_date$positions[reports_date$labels=="2021-02-23"] +15
```

```{r function_for_graph_self_assessement_bias}

g_self_assessment_bias <- function(
    error_df, error_metric, y_low_annotation, y_top_annotation, y_lim){
  
  #prepare data
  temp <- error_df %>%
    select(scenario_ID, error = !!as.symbol(error_metric), scenario_type, `Self-assessment by modelers`) %>%
    group_by(scenario_ID, scenario_type, `Self-assessment by modelers`) %>% #when both endpoint reports, take the mean of the 2
    summarise(
      error = mean(error)
      ) %>%
    mutate(
      scenario_ID = gsub("Scenario: ", "", scenario_ID)
    ) %>%
    spread(
      scenario_type, error
    )
    
  #main graph
  pmain <- ggplot(temp) +
    #dates of reports on x axis
    geom_rug(aes(as.Date(scenario_ID))) +
    #error line and points
    geom_pointrange(
      aes(
        y=median, ymin=`most\noptimistic`, ymax=`most\npessimistic`, x=as.Date(scenario_ID), 
        color=`Self-assessment by modelers`
        ),
      size=.4, linewidth=1, shape=16
      ) +
    #colors assessed by modelers or not
    scale_color_manual(
      values = c(
        "Yes" = alpha("#1e8449", 1), 
        "No" = alpha("#900C3F", .2)
        )
      ) +
    #arrow and accuracy annotation
    annotate(
      'segment', arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
      x = as.Date("2022-04-01"), xend = as.Date("2022-04-01"),
      y = y_low_annotation,  yend = y_top_annotation, alpha=.7,
      ) +
    annotate(
      "text", label="accurate", family="Times New Roman", fontface="italic",
      x=as.Date("2022-04-01"),
      y=y_low_annotation, vjust=1,
    ) +
    annotate(
      "text", label="unaccurate", family="Times New Roman", fontface="italic",
      x=as.Date("2022-04-01"),
      y=y_top_annotation, vjust=-0.3
    ) +
    #graph parameters
    scale_y_continuous(
      labels = scales::percent #% notation on y axis
      ) + 
    scale_x_continuous(#labels of reports
      breaks = reports_date$positions,
      labels = reports_date$labels
    ) +
    coord_cartesian(
      xlim = c(as.Date("2020-04-20"), as.Date("2022-05-01")),
      ylim = c(0, y_lim)
      ) +
    theme(
      axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
      axis.ticks.x=element_blank(),
      legend.position = c(.2, .7),
      legend.background = element_rect(fill=alpha('white', .7), color="grey")
      ) +
    labs(
      x="",
      y="Mean Absolute Error",
      color="scenarios\nself-assessed\nby modelers?"
      ) 
    
    #side boxplot
    ydens <- axis_canvas(pmain, axis = "y", coord_flip = TRUE) +
      geom_boxplot(
        data = temp, 
        aes(x = median, fill=`Self-assessment by modelers`),
        alpha = 0.5, size = 0.2, linewidth=.5, outlier.shape = 4
        ) + 
      xlim(0, y_lim) +
      scale_fill_manual(
        values = c(
          "Yes" = alpha("#1e8449", 1), 
          "No" = alpha("#900C3F", .2)
          )
        ) +
      coord_flip() 
  
  return(list(pmain, ydens))
  
}
```

### MAE

attention à avril 2020 =\> normalement OK, checker avec max (quel max a-t-il pris ? devrait être 2668)

rajouter celui des cas

probleme il pas noramisé que MAE optimist et pessimist, alors que le median oui !!

```{r}
temp <- read_csv('results/results_with_illegtimate_comparisons.csv') %>%
  rename(
    scenario_ID = Date,
    endpoints = Endpoint
  ) %>%
  #normalise MAE
  mutate(
    across(
      c(`MAE (optimist)`, `MAE (pessimist)`),
      ~case_when(
        endpoints=="ICU"~.x/max_ICU_beds,
        endpoints=="New hosp."~.x/max_new_hosp
      )
      )
    ) %>%
  #for % MAPE on graph
  mutate(
    across(
      c(`MAPE (median)`, `MAPE (optimist)`, `MAPE (pessimist)`, `MAE (median)`),
      ~.x/100
      )
    ) %>%
  gather(
    error_scenario, percent, `MAPE (median)`:`MAE (pessimist)`
    ) %>%
  separate(error_scenario, sep = " ", into=c("error_metric", "scenario_type")) %>%
  spread(error_metric, percent)

temp$scenario_type = factor(temp$scenario_type)
temp$scenario_type = fct_recode(
  temp$scenario_type,
  median = "(median)", `most\noptimistic` = "(optimist)", `most\npessimistic` = "(pessimist)"
  )
errors$scenario_type <- factor(
  errors$scenario_type,
  levels = c(
    'most\npessimistic',
    'median',
    'most\noptimistic'
  )
)

temp <- temp %>%
  rename(
    `Mean Absolute Error (%)` = MAE,
    `Mean Absolute Percentage Error` = MAPE
    )
  
```


```{r self_assessment_bias_vs_MAE}
g <- insert_yaxis_grob(
  g_self_assessment_bias(temp, "Mean Absolute Error (%)", .05, 0.7, NA)[[1]] +
    labs(y="Mean Absolute Error\n(% of historical peak)"), 
  g_self_assessment_bias(temp, "Mean Absolute Error (%)", .05, 0.7, NA)[[2]], 
  grid::unit(.1, "null"), position = "right"
  )
(ggdraw(g))

f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/self_assessment_bias_MAE",
  6, 4, 500
)
rm(g)
```

### MAPE

```{r self_assessment_bias_vs_MAPE}
g2 <- insert_yaxis_grob(
  g_self_assessment_bias(temp, "Mean Absolute Percentage Error", .2, 2.7, 3)[[1]] +
    labs(y="Mean Absolute Percentage Error"), 
  g_self_assessment_bias(temp, "Mean Absolute Percentage Error", .2, 2.7, 3)[[2]], 
  grid::unit(.1, "null"), position = "right"
  )
(ggdraw(g2))

f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/self_assessment_bias_MAPE",
  6, 4, 500
)
```

## Both graphs

```{r both_biases}
plot_grid(
  g1, g2, nrow=2, rel_heights = c(.45, .55), 
  labels = c("a", "b"), axis="l", align="v"
  )

f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/biases",
  6, 6, 500
)

rm(reports_date, g1, g2)
```

# Uncertainty (mean and max)

A faire : rajouter en second axis l'expression en % du peak (je pense que ça tient mathématiquement)

```{r uncertainties}
temp <- temporal %>%
  select(scenario_ID, endpoint, Period, `Average uncertainty (beds)`) 

x_arrow <- 4.7
x_label_uncertain <- 4.6
x_label_peak <- .7
y_arrow_ICU <- 6000
y_arrow_hosp <- 2500

hline_dat = data.frame(
  endpoint=c("ICU", "New hosp."),
  beds=c(max_ICU_beds, max_new_hosp)
  )

vline_dat = data.frame(
  endpoint=c("ICU", "New hosp."),
  x=c(x_arrow, x_arrow),
  xend=c(x_arrow, x_arrow),
  y=c(0, 0),
  yend=c(y_arrow_ICU, y_arrow_hosp)
  )

label_dat = data.frame(
  label=c("more\nuncertainty", "historical epidemic peak\n(April 2020)"),
  endpoint=c("ICU", "ICU"),
  x=c(x_label_uncertain, x_label_peak),
  y=c(y_arrow_ICU, max_ICU_beds)
  )

ggplot(temp) +
  geom_boxplot(
    aes(Period, `Average uncertainty (beds)`, fill=endpoint),
    position = position_nudge(y = 0, x = -.3),
    outlier.shape = NA,
    width=.3, alpha=.6
    ) +
  geom_quasirandom(
    aes(Period, `Average uncertainty (beds)`, color=endpoint),
    alpha=1, width=.1
    ) +
  labs(
    x='time since report publication', 
    y='',
    fill="report uncertainty\n(beds)",
    color="report uncertainty\n(beds)"
    ) +
  guides(
    fill="none",
    color = guide_legend(override.aes = list(size = 3))
    ) +
  scale_y_continuous(
    labels = scales::label_number(drop0trailing = TRUE)
  ) +
  theme(
    axis.line.x=element_blank(),
    strip.text.y = element_blank(),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
    legend.position = "left"
    ) +
  # annotate(
  #   "text", x = 3, y = 1, label = "historical maximum\nbed occupancy peak", hjust=1,
  #   fontface = "italic", family = "Times New Roman"
  #   ) +
  facet_grid(vars(endpoint), scales="free_y") +
  geom_hline(
    data=hline_dat, 
    aes(yintercept=beds), 
    linetype="dashed", alpha=.7
    ) +
  geom_hline(
    yintercept=0, 
    ) +
  geom_segment(
    data=vline_dat,
    aes(x=x, xend=xend, y=y, yend=yend), 
    arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
    alpha=.6
    ) +
  geom_text(
    data=label_dat,
    aes(x=x, y=y, label=label, vjust=c(1, 0.5), hjust=c(1, 0)), 
    fontface = "italic", family = "Times New Roman",
    ) +
  coord_cartesian(
    xlim = c(NA, 4.2)
    )

f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/uncertainties",
  8, 6, 500
)

rm(temp, hline_dat, vline_dat, x_arrow, x_label, x_label_peak, x_label_uncertain, y_arrow_hosp, y_arrow_hosp, label_dat)

#old code without temporal evolution 

# temp <- uncertainties %>%
#   
#   select(scenario_ID,`Average uncertainty`, `Max uncertainty`) %>%
#   group_by(scenario_ID) %>% #when both endpoint reports, take hte mean of the 2
#   summarise(
#     `Average uncertainty` = mean(`Average uncertainty`),
#     `Max uncertainty` = mean(`Max uncertainty`),
#     ) %>%
#   rename(
#     `Average\nUncertainty` = `Average uncertainty`, 
#     `Maximum\nUncertainty` = `Max uncertainty`, 
#     ) %>%
#   gather(
#     key=indicator, value=percentage, `Average\nUncertainty`, `Maximum\nUncertainty`
#     ) %>%
#   mutate(
#     percentage = percentage/100 #to have percentages on graph
#   )
# 
# ggplot(temp) +
#   geom_boxplot(
#     aes(indicator, percentage),
#     position = position_nudge(y = 0, x = -.3),
#     outlier.shape = NA,
#     width=.2, alpha=.6
#     ) +
#   geom_quasirandom(
#     aes(indicator, percentage),
#     alpha=.6, width=.1
#     ) +
#   scale_y_continuous(
#     labels = scales::percent,
#     breaks=c(seq(0, max(temp$percentage), .25))
#     ) +
#   labs(
#     x='', 
#     y='reports uncertainty'
#     ) +
#   theme(legend.position = 'none') +
#   geom_hline(yintercept=1, linetype='dashed') +
#   annotate(
#     "text", x = 3, y = 1, label = "historical maximum\nbed occupancy peak", hjust=1,
#     fontface = "italic", family = "Times New Roman"
#     )
```

# uncertainty vs error {.tabset .tabset-fade .tabset-pills}

## ICU

```{r}
temp <- temporal %>%
  gather(
    scenario_type, value=error, `MAE (high, beds)`, `MAE (low, beds)`, `MAE (median, beds)`
  ) %>%
  mutate(
    scenario_type=case_when(
      scenario_type == "MAE (high, beds)" ~ "pessimist",
      scenario_type == "MAE (median, beds)" ~ "median",
      scenario_type == "MAE (low, beds)" ~ "optimist"
    )
  ) %>%
  filter(endpoint=="ICU")

temp$scenario_type <- factor(
  temp$scenario_type,
  levels = c(
    'pessimist',
    'median',
    'optimist'
  )
)

axis_lim <- max(
  max(temp$error), temp$`Average uncertainty (beds)`, max_ICU_beds
  )

ggplot(temp) +
  geom_point(aes(error, `Average uncertainty (beds)`, color=scenario_type)) +
  facet_grid(endpoint~Period, scales="free_y") +
  ylim(0, axis_lim) +
  xlim(0, axis_lim) +
  geom_vline(xintercept = max_new_hosp, linetype="dashed") +
  geom_hline(yintercept = max_new_hosp, linetype="dashed") +
  coord_flip() +
  scale_color_viridis_d() +
  geom_line(aes(error, `Average uncertainty (beds)`, group=scenario_ID), alpha=.4)

```

## new hosp

```{r}
temp <- temporal %>%
  gather(
    scenario_type, value=error, `MAE (high, beds)`, `MAE (low, beds)`, `MAE (median, beds)`
  ) %>%
  mutate(
    scenario_type=case_when(
      scenario_type == "MAE (high, beds)" ~ "pessimist",
      scenario_type == "MAE (median, beds)" ~ "median",
      scenario_type == "MAE (low, beds)" ~ "optimist"
    )
  ) %>%
  filter(endpoint=="New hosp.")

temp$scenario_type <- factor(
  temp$scenario_type,
  levels = c(
    'pessimist',
    'median',
    'optimist'
  )
)

axis_lim <- max(
  max(temp$error), temp$`Average uncertainty (beds)`, max_new_hosp
  )

ggplot(temp) +
  geom_point(aes(error, `Average uncertainty (beds)`, color=scenario_type)) +
  facet_grid(endpoint~Period, scales="free_y") +
  ylim(0, axis_lim) +
  xlim(0, axis_lim) +
  geom_vline(xintercept = max_new_hosp, linetype="dashed") +
  geom_hline(yintercept = max_new_hosp, linetype="dashed") +
  coord_flip() +
  scale_color_viridis_d() +
  geom_line(aes(error, `Average uncertainty (beds)`, group=scenario_ID), alpha=.4)
```

## all together

# Error temporal

A affiner, pour l'instant je n'ai que le MAE median

si on avait le MAE du worst case scenario, on verrait que ça dépasse les 50% et s'approche des 100% souvent

ATTENTION QUAND C'EST EN ABSOLU IL FAUT ENLEVER LE SCENARIO D'AVRIL OU LUI APPLIQUER UNE TRANSFORMATION PROPORTIONNELLE ICU_FRANCE/ICU_IDF

```{r}
temp <- temporal %>%
  select(
    scenario_ID, endpoint, Period, 
    `MAPE (median)` = `MAE (median, beds)`, 
    `MAPE (most\npessimistic)` = `MAE (high, beds)`, 
    `MAPE (most\noptimistic)` = `MAE (low, beds)`
    ) %>%
  gather(
    key=indicator, value=beds, `MAPE (median)`, `MAPE (most\npessimistic)`, `MAPE (most\noptimistic)`
    ) %>%
  separate(indicator, sep = " ", into=c("error_metric", "scenario_type")) 

temp$scenario_type = factor(temp$scenario_type)
temp$scenario_type = fct_recode(
  temp$scenario_type,
  median = "(median)", `most\noptimistic` = "(most\noptimistic)", `most\npessimistic` = "(most\npessimistic)"
  )
temp$scenario_type <- factor(
  temp$scenario_type,
  levels = c(
    "most\npessimistic",
    "median",
    "most\noptimistic"
  )
)

hline_dat = data.frame(
  endpoint=c("ICU", "New hosp."),
  beds=c(max_ICU_beds, max_new_hosp)
  )

label_dat = data.frame(
  label=c("", "", "historical epidemic peak\n(April 2020)"),
  endpoint=c("ICU", "ICU", "ICU"),
  scenario_type=unique(temp$scenario_type),
  x=c(2, 2, 2),
  y=c(max_ICU_beds, max_ICU_beds, max_ICU_beds)
  )

ggplot(temp) +
  geom_boxplot(
    aes(Period, beds, fill=scenario_type), 
    position = position_nudge(y = 0, x = -.3),
    outlier.shape = NA,
    width=.3, alpha=.6
    ) +
  geom_quasirandom(
    aes(Period, beds, color=scenario_type),
    alpha=.6, width=.1
    ) +
  facet_grid(endpoint~scenario_type, scales="free_y") +
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  geom_hline(
    data=hline_dat, 
    aes(yintercept=beds), 
    linetype="dashed", alpha=.7
    ) +
  geom_hline(
    yintercept=0, 
    ) +
  scale_y_continuous(
    labels = scales::label_number(drop0trailing = TRUE)
  ) +
  theme(
    axis.line.x=element_blank(),
    strip.text.x = element_blank(),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)
    ) +
    labs(
      y="Mean Absolute Error\n(beds)",
      x="time since report publication",
      fill="in each report,\nscenario...",
      color="in each report,\nscenario..."
    ) +
  guides(
    fill="none",
    color = guide_legend(override.aes = list(size = 3))
    ) +
  geom_text(
    data=label_dat,
    aes(x=x, y=y, label=label),
    fontface = "italic", family = "Times New Roman",
    )

f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/MAE_time",
  8, 6, 500
)

rm(temp, hline_dat, label_dat)
```
