---
title: "Accuracy, Uncertainty, Bias"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

This pages plots the graphs of the main results concerning accuracy, uncertainty and bias of the scenarios reports.

```{r setup, include=T, message=F, warning=F}
knitr::opts_chunk$set(message=F, warning=F, fig.align = "center",  dev='png')

library(tidyverse) #loads multiple packages (see https://tidyverse.tidyverse.org/)

#core tidyverse packages loaded:
# ggplot2, for data visualisation. https://ggplot2.tidyverse.org/
# dplyr, for data manipulation. https://dplyr.tidyverse.org/
# tidyr, for data tidying. https://tidyr.tidyverse.org/
# readr, for data import. https://readr.tidyverse.org/
# purrr, for functional programming. https://purrr.tidyverse.org/
# tibble, for tibbles, a modern re-imagining of data frames. https://tibble.tidyverse.org/
# stringr, for strings. https://stringr.tidyverse.org/
# forcats, for factors. https://forcats.tidyverse.org/
# lubridate, for date/times. https://lubridate.tidyverse.org/

#also loads the following packages (less frequently used):
# Working with specific types of vectors:
#     hms, for times. https://hms.tidyverse.org/
# Importing other types of data:
#     feather, for sharing with Python and other languages. https://github.com/wesm/feather
#     haven, for SPSS, SAS and Stata files. https://haven.tidyverse.org/
#     httr, for web apis. https://httr.r-lib.org/
#     jsonlite for JSON. https://arxiv.org/abs/1403.2805
#     readxl, for .xls and .xlsx files. https://readxl.tidyverse.org/
#     rvest, for web scraping. https://rvest.tidyverse.org/
#     xml2, for XML. https://xml2.r-lib.org/
# Modelling
#     modelr, for modelling within a pipeline. https://modelr.tidyverse.org/
#     broom, for turning models into tidy data. https://broom.tidymodels.org/

# Use the conflicted package (<http://conflicted.r-lib.org/>) to force all conflicts to become errors


library(cowplot) #for plot_grid()
library(ggbeeswarm) #for geom_beeswarm()
library(ggnewscale) #to reset manual scale in ggplot when multiple fill scales

#set theme for graphs
theme_set(
  theme_classic() +
  theme(
    panel.grid.major.y = element_line(), #no vertical lines by default
    text = element_text(family = "Times New Roman"), #default font
    plot.title = element_text(face="bold"), #graphs titles in bolds
    )
  )

# epidemic peaks during 1st wave
max_new_hosp <- 3036 #historical peak of new hospitalizations
max_ICU_beds <- 6937 #historical peak of ICU beds
max_new_hosp_IDF <- 1210
max_ICU_beds_IDF <- 2852 #historical peak of ICU beds in ile de france region (for Apr 2020 scenario)
dpi_resolution <- 500 #default dpi resolution for our png graphs

# Get the directory of the R Markdown file and set the working directory to the file location
md_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(md_dir)
```

# Prepare data

We load the files generated from the jupyter notebooks, which computed all the errors metrics (mean absolute error, mean error, mean absolute percentage error, uncertainty...).

```{r load_data}
errors <-
  read_csv('results/error_metrics_stratified_by_scenario_types.csv') %>%
  mutate(
    #to have percentages on graphs
    MAE = MAE/100,
    ME = ME/100,
    `Max Error` = `Max Error`/100,
    MAPE = MAPE/100
    ) %>%
  rename(
    #for more explicit column names
    scenario_ID = ...1,
    scenario_type = scenario,
    `Mean Absolute Error (%)` = MAE,
    `Mean Error (%)` = ME,
    `Maximum Error (%)` = `Max Error`,
    `Mean Absolute Percentage Error` = MAPE,
    `Mean Absolute Error (beds)` = `MAE (beds)`,
    `Maximum Error (beds)` = `Max error (beds)`
    ) %>%
  mutate(
    #report ID with date of publication
    scenario_ID = as.Date(gsub("Scenario: ", "", scenario_ID)),
    #more explicit scenarios denomination
    scenario_type = case_when(
      scenario_type == "Optimist" ~ "best case",
      scenario_type == "Pessimist" ~ "worst case",
      scenario_type == "Median" ~ "median"
      ),
    #more explicit endpoint name
    endpoints = case_when(
      endpoints=="ICU"~"Intensive Care Units",
      endpoints=="New hosp."~"New Hospitalizations"
      )
    )

#order scenarios levels for graph order
errors$scenario_type <- factor(
  errors$scenario_type,
  levels = c(
    'worst case',
    'median',
    'best case'
  )
)

temporal <- read_csv('results/error_metrics_stratified_by_dates.csv') %>%
  select(-...1) %>%
  rename(
    scenario_ID = Scenario,
    endpoint = `Scenario type`
    ) %>%
  mutate(
    scenario_ID = as.Date(gsub("Scenario: ", "", scenario_ID)),
    endpoint = case_when(
      endpoint=="ICU"~"Intensive Care Units",
      endpoint=="New hosp."~"New Hospitalizations"
      )
  ) %>%
  #change time point names from days to weeks
  filter(Period!="56 days - 70 days") %>% #to few points for this period
  mutate(
    Period = case_when(
      Period == "0 days - 14 days" ~ "0-2 weeks",
      Period == "14 days - 28 days" ~ "2-4 weeks",
      Period == "28 days - 42 days" ~ "4-6 weeks",
      Period == "42 days - 56 days" ~ "6-8 weeks",
      Period == "" ~ ""
    )
  )
```

For April 2020 scenarios which only concern Ile-de-France region, the absolute errors and uncertainty expressed in number of beds are not directly comparable to the national figure. To account of this, in that particular case, we apply a proportional factor to account for the fact that number of beds is lower than at the national scale. This proportional factor is the ratio between the historical epidemic peak at the national scale and the one at the regional Ile-de-France scale, or a `r round(max_ICU_beds/max_ICU_beds_IDF, 1)` factor.

```{r correct_IDF_data}
temporal$`Average uncertainty (beds)`[temporal$scenario_ID=="2020-04-28"] <- 
  temporal$`Average uncertainty (beds)`[temporal$scenario_ID=="2020-04-28"]*max_ICU_beds/max_ICU_beds_IDF

temporal$`MAE (low, beds)`[temporal$scenario_ID=="2020-04-28"] <- 
  temporal$`MAE (low, beds)`[temporal$scenario_ID=="2020-04-28"]*max_ICU_beds/max_ICU_beds_IDF

temporal$`MAE (median, beds)`[temporal$scenario_ID=="2020-04-28"] <- 
  temporal$`MAE (median, beds)`[temporal$scenario_ID=="2020-04-28"]*max_ICU_beds/max_ICU_beds_IDF

temporal$`MAE (high, beds)`[temporal$scenario_ID=="2020-04-28"] <- 
  temporal$`MAE (high, beds)`[temporal$scenario_ID=="2020-04-28"]*max_ICU_beds/max_ICU_beds_IDF
```

```{r function_save_pdf_ong}
#function to save plots as pdf and png
f_save_graph_pdf_png <- function(path_name, graph_width, graph_height, dpi_resolution){
  #pdf
  ggsave(
    paste0(path_name, ".pdf"),
    width=graph_width, height=graph_height, bg="white", 
    device = cairo_pdf #devide cairo for Times New Roman font in pdf
    )
  #png
  ggsave(
    paste0(path_name, ".png"),
    width=graph_width, height=graph_height, bg="white", 
    dpi = dpi_resolution
    )
}
```

# Absolute Error by 2-weeks period

```{r}
temp <- temporal %>%
  select(
    scenario_ID, endpoint, Period, 
    `MAE (median)` = `MAE (median, beds)`, 
    `MAE (worst\ncase)` = `MAE (high, beds)`, 
    `MAE (best\ncase)` = `MAE (low, beds)`
    ) %>%
  gather(
    key=indicator, value=beds, `MAE (median)`, `MAE (worst\ncase)`, `MAE (best\ncase)`
    ) %>%
  separate(indicator, sep = " ", into=c("error_metric", "scenario_type")) 

temp$scenario_type = factor(temp$scenario_type)
temp$scenario_type = fct_recode(
  temp$scenario_type,
  median = "(median)", `best case` = "(best\ncase)", `worst case` = "(worst\ncase)"
  )
temp$scenario_type <- factor(
  temp$scenario_type,
  levels = c(
    "worst case",
    "median",
    "best case"
  )
)

hline_dat = data.frame(
  endpoint=c("Intensive Care Units", "New Hospitalizations"),
  beds=c(max_ICU_beds, max_new_hosp)
  )

label_dat = data.frame(
  label=c("historical epidemic peak\n(April 2020)"),
  endpoint=c("Intensive Care Units"),
  scenario_type=c("best case"),
  x=c(4),
  y=c(max_ICU_beds)
  )

ggplot(temp) +
  geom_boxplot(
    aes(Period, beds, fill=scenario_type), 
    position = position_nudge(y = 0, x = -.3),
    outlier.shape = NA,
    width=.3, alpha=.6
    ) +
  geom_quasirandom(
    aes(Period, beds, color=scenario_type),
    alpha=.6, width=.1
    ) +
  facet_grid(endpoint~scenario_type, scales="free_y") +
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  geom_hline(
    data=hline_dat, 
    aes(yintercept=beds), 
    linetype="dashed", alpha=.7
    ) +
  geom_hline(
    yintercept=0, 
    ) +
  scale_y_continuous(
    labels = scales::label_number(drop0trailing = TRUE)
  ) +
  theme(
    axis.line.x=element_blank(),
    strip.text.x = element_blank(),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)
    ) +
    labs(
      y="Mean Absolute Error\n(beds)",
      x="time since report publication",
      fill="in each report,\nscenario...",
      color="in each report,\nscenario..."
    ) +
  guides(
    fill="none",
    color = guide_legend(override.aes = list(size = 3))
    ) +
  geom_text(
    data=label_dat,
    aes(x=x, y=y, label=label), hjust=1,
    fontface = "italic", family = "Times New Roman",
    )

f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/MAE_time",
  8, 6, 500
)

rm(temp, hline_dat, label_dat)
```


# Bias {.tabset .tabset-fade .tabset-pills}

## Error Bias (Mean Error) {.tabset .tabset-fade}

When a report has scenarios relateing to both ICU and new hospitalizations, we take the mean of the 2.

```{r bias_mean_error}
temp <- errors %>%
  select(scenario_ID, scenario_type, `Mean Error (%)`) %>%
  group_by(scenario_ID, scenario_type) %>%
  summarise(
    `Mean Error (%)` = mean(`Mean Error (%)`)
    )
temp$scenario_type <- fct_recode(
  temp$scenario_type, 
  `worst case scenarios\nof reports` = "worst case", 
  `median scenarios\nof reports` = "median", 
  `best case scenarios\nof reports` = "best case"
  )

(g1 <- ggplot(temp) +
    geom_boxplot(
      aes(scenario_type, `Mean Error (%)`, fill=scenario_type),
      position = position_nudge(y = 0, x = -.3),
      outlier.shape = NA,
      width=.2, alpha=.6
      ) +
    geom_quasirandom(
        aes(scenario_type, `Mean Error (%)`, color=scenario_type),
        alpha=.6, width=.1
        ) +
    scale_y_continuous(
      labels = scales::percent,
      limits=c(-.8, .8),
      breaks = seq(-.75, .75, .25)
      ) +
    labs(
      x='',
      y='Mean Error'
      ) +
    scale_fill_viridis_d() +
    scale_color_viridis_d() +
    theme(legend.position = 'none') +
    geom_hline(yintercept=0, linetype='dashed') +
    annotate(
      'segment', arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
      x = 3.1, y = 0.1, xend = 3.1, yend = .7,  alpha=.5
      ) +
    annotate(
      'segment', arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
      x = 3.1, y = -0.1, xend = 3.1, yend = -.7, alpha=.5
      ) +
    annotate(
      "text", x = 3.15, y = .1, label = "unbiased", hjust=0, 
      fontface = "italic", family = "Times New Roman",
      ) +
    annotate(
      "text", x = 3, y = -.7, label = "too\noptimistic", hjust=1, vjust=0,
      fontface = "italic", family = "Times New Roman",
      ) +
    annotate(
      "text", x = 3, y = .7, label = "too\npessimistic", hjust=1, vjust=1,
      fontface = "italic", family = "Times New Roman",
      )
)


f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/bias_mean_error",
  6, 4, 500
)
```

## Self-assessment bias {.tabset .tabset-fade}

```{r tweak_reports_labels_for_better_display_on_graph}
#small offset to position adjacent reports labels on x axis
reports_date <-  data.frame(
  positions = unique(as.Date(gsub("Scenario: ", "", errors$scenario_ID))),
  labels = unique(as.Date(gsub("Scenario: ", "", errors$scenario_ID)))
)
#Aug and Jul 2021
reports_date$positions[reports_date$labels=="2021/07/26"] <- reports_date$positions[reports_date$labels=="2021/07/26"] -10
reports_date$positions[reports_date$labels=="2021/08/05"] <- reports_date$positions[reports_date$labels=="2021/08/05"] +10
#Apr and May 2021
reports_date$positions[reports_date$labels=="2021-04-26"] <- reports_date$positions[reports_date$labels=="2021-04-26"] -5
reports_date$positions[reports_date$labels=="2021-05-21"] <- reports_date$positions[reports_date$labels=="2021-05-21"] +5
#Jan and Feb 2021
reports_date$positions[reports_date$labels=="2021-01-16"] <- reports_date$positions[reports_date$labels=="2021-01-16"] -20
reports_date$positions[reports_date$labels=="2021-02-02"] <- reports_date$positions[reports_date$labels=="2021-02-02"] -10
reports_date$positions[reports_date$labels=="2021-02-08"] <- reports_date$positions[reports_date$labels=="2021-02-08"] +5
reports_date$positions[reports_date$labels=="2021-02-23"] <- reports_date$positions[reports_date$labels=="2021-02-23"] +15
```

```{r function_for_graph_self_assessement_bias}
g_self_assessment_bias <- function(
    error_df, error_metric, y_low_annotation, y_top_annotation, y_lim){
  
  #prepare data
  temp <- error_df %>%
    select(scenario_ID, error = !!as.symbol(error_metric), scenario_type, `Self-assessment by modelers`) %>%
    group_by(scenario_ID, scenario_type, `Self-assessment by modelers`) %>% #when both endpoint reports, take the mean of the 2
    summarise(
      error = mean(error)
      ) %>%
    mutate(
      scenario_ID = gsub("Scenario: ", "", scenario_ID)
    ) %>%
    spread(
      scenario_type, error
    )
    
  #main graph
  pmain <- ggplot(temp) +
    #dates of reports on x axis
    geom_rug(aes(as.Date(scenario_ID))) +
    #error line and points
    geom_pointrange(
      aes(
        y=median, ymin=`best case`, ymax=`worst case`, x=as.Date(scenario_ID), 
        color=`Self-assessment by modelers`
        ),
      size=.4, linewidth=1, shape=16
      ) +
    #colors assessed by modelers or not
    scale_color_manual(
      values = c(
        "Yes" = alpha("#1e8449", 1), 
        "No" = alpha("#900C3F", .2)
        )
      ) +
    #arrow and accuracy annotation
    annotate(
      'segment', arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
      x = as.Date("2022-04-01"), xend = as.Date("2022-04-01"),
      y = y_low_annotation,  yend = y_top_annotation, alpha=.7,
      ) +
    annotate(
      "text", label="accurate", family="Times New Roman", fontface="italic",
      x=as.Date("2022-04-01"),
      y=y_low_annotation, vjust=1,
    ) +
    annotate(
      "text", label="unaccurate", family="Times New Roman", fontface="italic",
      x=as.Date("2022-04-01"),
      y=y_top_annotation, vjust=-0.3
    ) +
    #graph parameters
    scale_y_continuous(
      labels = scales::percent #% notation on y axis
      ) + 
    scale_x_continuous(#labels of reports
      breaks = reports_date$positions,
      labels = format(reports_date$labels, format="%b %d, %Y")
    ) +
    coord_cartesian(
      xlim = c(as.Date("2020-04-20"), as.Date("2022-05-01")),
      ylim = c(0, y_lim)
      ) +
    theme(
      axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
      axis.ticks.x=element_blank(),
      legend.position = c(.2, .7),
      legend.background = element_rect(fill=alpha('white', .7), color="grey")
      ) +
    labs(
      x="",
      y="Mean Absolute Error",
      color="scenarios\nself-assessed\nby modelers?"
      ) 
    
    #side boxplot
    ydens <- axis_canvas(pmain, axis = "y", coord_flip = TRUE) +
      geom_boxplot(
        data = temp, 
        aes(x = median, fill=`Self-assessment by modelers`),
        alpha = 0.5, size = 0.2, linewidth=.5, outlier.shape = 4
        ) + 
      xlim(0, y_lim) +
      scale_fill_manual(
        values = c(
          "Yes" = alpha("#1e8449", 1), 
          "No" = alpha("#900C3F", .2)
          )
        ) +
      coord_flip() 
  
  return(list(pmain, ydens))
  
}
```

```{r}
temp <- read_csv('results/error_metrics_including_illegitimate_comparisons.csv') %>%
  rename(
    scenario_ID = Date,
    endpoints = Endpoint
  ) %>%
  #for % on graph
  mutate(
    across(
      c(`MAPE (median)`, `MAPE (optimist)`, `MAPE (pessimist)`, `MAE (median)`, `MAE (optimist)`, `MAE (pessimist)`, ),
      ~.x/100
      )
    ) %>%
  gather(
    error_scenario, percent, `MAE (median)`:`MAPE (pessimist)`
    ) %>%
  separate(error_scenario, sep = " ", into=c("error_metric", "scenario_type")) %>%
  spread(error_metric, percent)

temp$scenario_type = factor(temp$scenario_type)
temp$scenario_type = fct_recode(
  temp$scenario_type,
  median = "(median)", `best case` = "(optimist)", `worst case` = "(pessimist)"
  )
errors$scenario_type <- factor(
  errors$scenario_type,
  levels = c(
    'worst case',
    'median',
    'best case'
  )
)

temp <- temp %>%
  rename(
    `Mean Absolute Error (%)` = MAE,
    `Mean Absolute Percentage Error` = MAPE
    )
  
```

### MAE

```{r self_assessment_bias_vs_MAE}
g <- insert_yaxis_grob(
  g_self_assessment_bias(temp, "Mean Absolute Error (%)", .05, 0.7, 1)[[1]] +
    labs(y="Mean Absolute Error\n(% of historical peak)"), 
  g_self_assessment_bias(temp, "Mean Absolute Error (%)", .05, 0.7, 1)[[2]], 
  grid::unit(.1, "null"), position = "right"
  )
(ggdraw(g))

f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/self_assessment_bias_MAE",
  6, 4, 500
)
rm(g)
```

### MAPE

```{r self_assessment_bias_vs_MAPE}
g2 <- insert_yaxis_grob(
  g_self_assessment_bias(temp, "Mean Absolute Percentage Error", .2, 2.7, 3)[[1]] +
    labs(y="Mean Absolute Percentage Error"), 
  g_self_assessment_bias(temp, "Mean Absolute Percentage Error", .2, 2.7, 3)[[2]], 
  grid::unit(.1, "null"), position = "right"
  )
(ggdraw(g2))

f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/self_assessment_bias_MAPE",
  6, 4, 500
)
```

## Both graphs

```{r both_biases}
plot_grid(
  g1, g2, nrow=2, rel_heights = c(.45, .55), 
  labels = c("a", "b"), axis="l", align="v"
  )

f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/biases",
  6, 6, 500
)

rm(reports_date, g1, g2)
```

# Uncertainty

```{r uncertainties}
temp <- temporal %>%
  select(scenario_ID, endpoint, Period, `Average uncertainty (beds)`) 

x_arrow <- 4.7
x_label_uncertain <- 4.6
x_label_peak <- .7
y_arrow_ICU <- 6000
y_arrow_hosp <- 2500

hline_dat = data.frame(
  endpoint=c("Intensive Care Units", "New Hospitalizations"),
  beds=c(max_ICU_beds, max_new_hosp)
  )

vline_dat = data.frame(
  endpoint=c("Intensive Care Units", "New Hospitalizations"),
  x=c(x_arrow, x_arrow),
  xend=c(x_arrow, x_arrow),
  y=c(0, 0),
  yend=c(y_arrow_ICU, y_arrow_hosp)
  )

label_dat = data.frame(
  label=c("more\nuncertainty", "historical epidemic peak\n(April 2020)"),
  endpoint=c("Intensive Care Units", "Intensive Care Units"),
  x=c(x_label_uncertain, x_label_peak),
  y=c(y_arrow_ICU, max_ICU_beds)
  )

ggplot(temp) +
  geom_boxplot(
    aes(Period, `Average uncertainty (beds)`, fill=endpoint),
    position = position_nudge(y = 0, x = -.3),
    outlier.shape = NA,
    width=.3, alpha=.6
    ) +
  geom_quasirandom(
    aes(Period, `Average uncertainty (beds)`, color=endpoint),
    alpha=1, width=.1
    ) +
  labs(
    x='time since report publication', 
    y='',
    fill="report uncertainty\n(beds)",
    color="report uncertainty\n(beds)"
    ) +
  guides(
    fill="none",
    color = guide_legend(override.aes = list(size = 3))
    ) +
  scale_y_continuous(
    labels = scales::label_number(drop0trailing = TRUE)
  ) +
  theme(
    axis.line.x=element_blank(),
    strip.text.y = element_blank(),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
    legend.position = "left"
    ) +
  # annotate(
  #   "text", x = 3, y = 1, label = "historical maximum\nbed occupancy peak", hjust=1,
  #   fontface = "italic", family = "Times New Roman"
  #   ) +
  facet_grid(vars(endpoint), scales="free_y") +
  geom_hline(
    data=hline_dat, 
    aes(yintercept=beds), 
    linetype="dashed", alpha=.7
    ) +
  geom_hline(
    yintercept=0, 
    ) +
  geom_segment(
    data=vline_dat,
    aes(x=x, xend=xend, y=y, yend=yend), 
    arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
    alpha=.6
    ) +
  geom_text(
    data=label_dat,
    aes(x=x, y=y, label=label, vjust=c(1, 0.5), hjust=c(1, 0)), 
    fontface = "italic", family = "Times New Roman",
    ) +
  coord_cartesian(
    xlim = c(NA, 4.2)
    )

f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/uncertainties",
  8, 6, 500
)

rm(temp, hline_dat, vline_dat, x_arrow, x_label, x_label_peak, x_label_uncertain, y_arrow_hosp, y_arrow_hosp, label_dat)

#old code without temporal evolution 

# temp <- uncertainties %>%
#   
#   select(scenario_ID,`Average uncertainty`, `Max uncertainty`) %>%
#   group_by(scenario_ID) %>% #when both endpoint reports, take hte mean of the 2
#   summarise(
#     `Average uncertainty` = mean(`Average uncertainty`),
#     `Max uncertainty` = mean(`Max uncertainty`),
#     ) %>%
#   rename(
#     `Average\nUncertainty` = `Average uncertainty`, 
#     `Maximum\nUncertainty` = `Max uncertainty`, 
#     ) %>%
#   gather(
#     key=indicator, value=percentage, `Average\nUncertainty`, `Maximum\nUncertainty`
#     ) %>%
#   mutate(
#     percentage = percentage/100 #to have percentages on graph
#   )
# 
# ggplot(temp) +
#   geom_boxplot(
#     aes(indicator, percentage),
#     position = position_nudge(y = 0, x = -.3),
#     outlier.shape = NA,
#     width=.2, alpha=.6
#     ) +
#   geom_quasirandom(
#     aes(indicator, percentage),
#     alpha=.6, width=.1
#     ) +
#   scale_y_continuous(
#     labels = scales::percent,
#     breaks=c(seq(0, max(temp$percentage), .25))
#     ) +
#   labs(
#     x='', 
#     y='reports uncertainty'
#     ) +
#   theme(legend.position = 'none') +
#   geom_hline(yintercept=1, linetype='dashed') +
#   annotate(
#     "text", x = 3, y = 1, label = "historical maximum\nbed occupancy peak", hjust=1,
#     fontface = "italic", family = "Times New Roman"
#     )
```

# Uncertainty vs Error {.tabset .tabset-fade .tabset-pills}

```{r}
temp <- temporal %>%
  gather(
    scenario_type, value=error, `MAE (high, beds)`, `MAE (low, beds)`, `MAE (median, beds)`
  ) %>%
  mutate(
    scenario_type=case_when(
      scenario_type == "MAE (high, beds)" ~ "worst case",
      scenario_type == "MAE (median, beds)" ~ "median",
      scenario_type == "MAE (low, beds)" ~ "best case"
    ),
    `Average uncertainty (beds)` = case_when(
      endpoint=="Intensive Care Units"~round(`Average uncertainty (beds)`/max_ICU_beds, 2),
      endpoint=="New Hospitalizations"~round(`Average uncertainty (beds)`/max_new_hosp, 2),
    ),
    error = case_when(
      endpoint=="Intensive Care Units"~round(error/max_ICU_beds, 2),
      endpoint=="New Hospitalizations"~round(error/max_new_hosp, 2),
    )
  )

#order scenarios types
  temp$scenario_type <- factor(
    temp$scenario_type,
    levels = c(
      'worst case',
      'median',
      'best case'
      )
    )

g_uncert_vs_error <- function(data, axis_lim, x_label_accurate, y_label_accurate){
  #data frame for background color
  x <- seq(0, axis_lim, 0.01)
  y <- seq(0, axis_lim, 0.01)
  d1 <- expand.grid(x = x, y = y)
  d1 <- d1 %>%
    rowwise() %>%
    mutate(
      color= sqrt((x/2)^2+y^2)
    )
  
  #labels (un)accurate and (un)certain
  label_dat = data.frame(
    label=c("unaccurate,\ncertain", "unaccurate,\nuncertain", "accurate,\ncertain", "accurate,\nuncertain"),
    Period=c("0-2 weeks", "0-2 weeks", "0-2 weeks", "0-2 weeks"),
    x=c(.01, axis_lim, x_label_accurate, axis_lim),
    y=c(axis_lim, axis_lim, y_label_accurate, .01),
    hjust=c(0, 1, 0, 1),
    vjust=c(1, 1, 0, 0)
    )
  
  
  g <- ggplot(data) +
    #the background gradient color
    geom_raster(
      data = d1,
      aes(x, y, fill = color),
      alpha=.5, show.legend = FALSE
      ) +
    scale_fill_gradientn(
      colors = c("green", "green", "orange", "orange", "red", "red", "purple", "purple"),
      values = c(0, .15, .3, .5,  .8, max(d1$color))/max(d1$color)
      ) +
    new_scale_fill() + #for new fill scale for points
    #the lines gathering scenarios of a same report
    geom_line(
      aes(`Average uncertainty (beds)`, error, group=interaction(scenario_ID,endpoint)), 
      alpha=.6, 
      ) +
    #points of scenarios
    geom_point(
      aes(`Average uncertainty (beds)`, error, fill=scenario_type), 
      shape=21, size=2
      ) +
    scale_fill_viridis_d() + 
    #the 4 period panes
    facet_wrap(vars(Period), scales="free_y") +
    #to have x and y axis intersect at 0 and % and +/- sign
    scale_x_continuous (limits=c(0,axis_lim), expand=c(0,0), labels = scales::percent) +
    scale_y_continuous(limits=c(0,axis_lim), expand=c(0,0), labels = scales::percent_format(prefix = "±")) +
    labs(
      x="Uncertainty of report\n(% of historical peak)",
      y="Mean Absolute Error of scenario\n(% of historical peak)",
      fill="scenario in report",
      subtitle = "scenarios on same vertical line are issued from same report"
    ) +
    #labels (un)accurate and (un)certain
    geom_text(
      data = label_dat,
      aes(x = x, y = y, label= label, hjust=hjust, vjust=vjust), 
        fontface = "italic", family = "Times New Roman",
        ) +
    #to have bigger points in legend
    guides(
      fill = guide_legend(override.aes = list(size = 3))
      ) +
    #no vertical lines by default, add them
    theme(
      panel.grid.major.x = element_line() 
    )
  
  return(g)
}
```


## ICU

```{r}
g_uncert_vs_error(
  temp %>% filter(endpoint=="Intensive Care Units"), 
  1.5, #axis limit
  .2, #x for label on bottom left
  0 #y for label on bottom left
  )
```

## new hosp

```{r}
g_uncert_vs_error(
  temp %>% filter(endpoint=="New Hospitalizations"), 
  1.2, #axis limit
  .3, #x for label on bottom left
  0 #y for label on bottom left
  )
```

## all together

When a report has scenarios relateing to both ICU and new hospitalizations, we take the mean of the 2.

```{r}
temp2 <- temp %>% 
  group_by(
    scenario_ID, scenario_type, Period
  ) %>%
  summarise(
    `Average uncertainty (beds)` = mean(`Average uncertainty (beds)`), 
    error = mean(error)
  ) %>%
  mutate(endpoint = "ICU and New Hosp")

g_uncert_vs_error(
  temp2, 
  1.5, #axis limit
  .3, #x for label on bottom left
  0 #y for label on bottom left
  )

f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/uncertainty_vs_error",
  7, 5, 500
)
```

# Other Absolute Errors {.tabset .tabset-fade .tabset-pills}

## MAE vs MAPE {.tabset .tabset-fade}

```{r function_graph_MAPE_vs_MAE}
#graph code adapted from http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/78-perfect-scatter-plots-with-correlation-and-marginal-histograms/

g_MAPE_vs_MAE <- function(df){
  # Main plot
  pmain <- ggplot(temp)+
    geom_point(
      aes(
        y = `Mean Absolute Error (%)`, x = `Mean Absolute Percentage Error`, 
        color = scenario_type, fill = scenario_type
        ),
      alpha = 0.6, stroke=1
      )+
    theme(
      panel.grid.major.x = element_line(),
      legend.position = c(.75, .4),
      legend.background = element_rect(fill=alpha('white', .5), color="grey")
      ) +
    scale_y_continuous(labels = scales::percent) +
    scale_x_continuous(
      labels=scales::percent(c(seq(0, max(temp$`Mean Absolute Percentage Error`), 1), .5)),
      breaks = c(seq(0, max(temp$`Mean Absolute Percentage Error`), 1), .5)
      ) +
    scale_color_viridis_d() +
    labs(
      y="Mean Absolute Error\n(% of historical bed occupancy peak)",
      x="Mean Absolute Percentage Error\n(% of real epidemic activity)",
      color="scenario from\neach report",
      fill="scenario from\neach report"
    ) +
    geom_hline(yintercept = 1, linetype="dashed", alpha=.4) 
  
  # Marginal densities along x axis
  xdens <- axis_canvas(pmain, axis = "x") +
    geom_boxplot(
      data = temp, 
      aes(x = `Mean Absolute Percentage Error`, fill=scenario_type),
      alpha = 0.5, size = 0.2, linewidth=.5, outlier.shape = 4
      ) +
    scale_fill_viridis_d()
  
  # Marginal densities along y axis
  ydens <- axis_canvas(pmain, axis = "y", coord_flip = TRUE) + # Need to set coord_flip = TRUE, if you plan to use coord_flip()
    geom_boxplot(
      data = temp, 
      aes(x = `Mean Absolute Error (%)`, fill = scenario_type),
      alpha = 0.5, size = 0.2, linewidth=.5, outlier.shape = 4
      ) +
    coord_flip() +
    scale_fill_viridis_d() 
  
  return(
    list(pmain, xdens,  ydens)
    )
}
```

### ICU + New Hosp

When a report has scenarios relateing to both ICU and new hospitalizations, we take the mean of the 2.

```{r MAPE_vs_MAE_ICU_new_hosp_combined}
#when scenarios relate to both ICU and new hosp, we take the mean of the 2
temp <- errors %>%
  select(
    scenario_ID, scenario_type, 
    `Mean Absolute Percentage Error`, `Mean Absolute Error (%)`
    ) %>%
  group_by(scenario_ID, scenario_type) %>%
  summarise(
    across(
      c(`Mean Absolute Percentage Error`, `Mean Absolute Error (%)`), 
      ~mean(.x)
      )
  )

#create plot
p1 <- insert_xaxis_grob(
  g_MAPE_vs_MAE(temp)[[1]], 
  g_MAPE_vs_MAE(temp)[[2]], 
  grid::unit(.2, "null"), position = "top"
  )
p2<- insert_yaxis_grob(
  p1, 
  g_MAPE_vs_MAE(temp)[[3]], 
  grid::unit(.2, "null"), position = "right"
  )
(g1 <- ggdraw(p2))

#save
f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/MAE_vs_MAPE",
  4, 4, dpi_resolution
)

rm(p1, p2, g1)

#same but with no legend (for subsequent graphs combination)
p1_no_legend <- insert_xaxis_grob(
  g_MAPE_vs_MAE(temp)[[1]] + theme(legend.position = "none"), 
  g_MAPE_vs_MAE(temp)[[2]], 
  grid::unit(.2, "null"), position = "top"
  )
p2_no_legend<- insert_yaxis_grob(
  p1_no_legend, 
  g_MAPE_vs_MAE(temp)[[3]], 
  grid::unit(.2, "null"), position = "right"
  )
g1_no_legend <- ggdraw(p2_no_legend)

rm(p1_no_legend, p2_no_legend, temp)
```

### ICU alone

```{r MAPE_vs_MAE_ICU_alone}
temp <- errors %>%
  filter(endpoints=="Intensive Care Units")

#create plot
p1 <- insert_xaxis_grob(
  g_MAPE_vs_MAE(temp)[[1]], 
  g_MAPE_vs_MAE(temp)[[2]], 
  grid::unit(.2, "null"), position = "top"
  )
p2<- insert_yaxis_grob(
  p1, 
  g_MAPE_vs_MAE(temp)[[3]], 
  grid::unit(.2, "null"), position = "right"
  )
(g1 <- ggdraw(p2))

rm(p1, p2, g1, temp)
```

### new hosp alone

```{r MAPE_vs_MAE_new_hosp_alone}
temp <- errors %>%
  filter(endpoints=="New Hospitalizations")

#create plot
p1 <- insert_xaxis_grob(
  g_MAPE_vs_MAE(temp)[[1]], 
  g_MAPE_vs_MAE(temp)[[2]], 
  grid::unit(.2, "null"), position = "top"
  )
p2<- insert_yaxis_grob(
  p1, 
  g_MAPE_vs_MAE(temp)[[3]], 
  grid::unit(.2, "null"), position = "right"
  )
(g1 <- ggdraw(p2))

rm(p1, p2, g1, temp)
```

## Absolute mean and max error (MAE)

```{r max_mean_absolute_error_beds}
temp <- errors %>%
  select(
    scenario_ID, 
    scenario_type, 
    endpoints,
    `Maximum\nAbsolute Error`=`Maximum Error (beds)`, 
    `Mean\nAbsolute Error`= `Mean Absolute Error (beds)`
    ) %>%
  gather(
    key=indicator, value=percentage, 
    `Maximum\nAbsolute Error`, `Mean\nAbsolute Error`
    )

hline_dat = data.frame(
  endpoints=c("Intensive Care Units", "New Hospitalizations"),
  beds=c(max_ICU_beds, max_new_hosp)
  )

(g2 <- ggplot(temp) +
    geom_boxplot(
      aes(scenario_type, percentage, fill=scenario_type),
      position = position_nudge(y = 0, x = -.35),
      outlier.shape = NA, width=.5, alpha=.6
      ) +
    geom_quasirandom(
        aes(scenario_type, percentage, color=scenario_type),
        alpha=.6, width=.1,
        ) +
    geom_hline(
      data=hline_dat, 
      aes(yintercept=beds), 
      linetype="dashed", alpha=.4
      ) +
    facet_grid(endpoints~indicator, scales = "free") +
    labs(
      x='', 
      y='error in terms of beds',
      color="scenarios from\neach report",
      fill="scenarios from\neach report",
      subtitle = "horizontal lines: historical bed occupancy peak"
      ) +
    guides(
      fill="none",
      color = guide_legend(override.aes = list(size = 3))
      ) +
    scale_fill_viridis_d() +
    scale_color_viridis_d() +
    scale_y_continuous(
      labels = scales::label_number(drop0trailing = TRUE)
    ) +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.line.x = element_blank()
    ) +
    geom_hline(yintercept = 0)
)

rm(temp, hline_dat)

#save
f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/mean_max_beds_error",
  6, 4, dpi_resolution
)
```

## Both graphs

```{r all_error_graphs}
plot_grid(
  g2, g1_no_legend ,
  rel_widths = c(5.5/9, 3.5/9),
  labels = c("a", "b")
)

#save
f_save_graph_pdf_png(
  "graphs/errors_and_uncertainties/errors",
  9, 3.5, dpi_resolution
)

rm(g1_no_legend, g2)
```

